### 2014-10-24 manuscript/paper-tycho.tex
Paper draft 10/24, faster adaptive errors
Paper draft sent around; adaptive error code steps more quickly
(better criteria for changing step size based on chi-sqr threshold)

%\documentclass[manuscript]{aastex}  % one-column, double-spaced GENERATE BIB
%\documentclass[12pt,preprint]{aastex}  % one-column, single-spaced
%\documentclass[iop, apj, numberedappendix, twocolappendix]{emulateapj}
\documentclass[iop, apj, numberedappendix]{emulateapj}

% Customize figures/sizing for 2 column vs. manuscript printout
\usepackage{etoolbox}
\newtoggle{manuscript}
%\toggletrue{manuscript}
\togglefalse{manuscript}

\shorttitle{Tycho's Synchrotron Rims (\today)}  % <~ 44 char
\shortauthors{XXX et al. (\today)}  % Max three
\slugcomment{Draft, \today}  % short title pg comment

%% ==================================================================== %%
%% README for track changes                                             %%
%% To add/remove text or add comments, use the following commands:      %%
%%                                                                      %%
%%       \note[editor]{The note}                                        %%
%%     \annote[editor]{Text to annotate}{The note}                      %%
%%        \add[editor]{Text to add}                                     %%
%%     \remove[editor]{Text to remove}                                  %%
%%     \change[editor]{Text to remove}{Text to add}                     %%
%%                                                                      %%
%% ==================================================================== %%

\usepackage[inline]{trackchanges}  % trackchanges.sourceforge.net
\addeditor{Rob}
\addeditor{Sean}
\addeditor{Steve}
\addeditor{Aaron}
\addeditor{Brian}

\usepackage{amsmath}  % amsthm, amssymb
\usepackage{CJK}  % aas.org/authors/author-names-non-roman-alphabets
\usepackage{booktabs}
\usepackage{hyperref}

\newcommand*{\mt}{\mathrm}
\newcommand*{\unit}[1]{\;\mt{#1}}  % vemod.net/typesetting-units-in-latex
\newcommand*{\abt}{\mathord{\sim}} % tex.stackexchange.com/q/55701
\newcommand*{\ptl}{\partial}
\newcommand*{\del}{\nabla}

% This paper
\newcommand*\mean[1]{\bar{#1}}
\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand*{\tsup}{\textsuperscript}

\defcitealias{ressler2014}{R14}
\newcommand*{\Chandra}{\textit{Chandra}\ }
\newcommand*{\tsynch}{\tau_{\mt{synch}}}
\newcommand*{\mE}{m_\mt{E}}
\newcommand*{\Ecut}{E_{\mt{cut}}}
\newcommand*{\Bmin}{B_{\mt{min}}}
\newcommand*{\muG}{\unit{\mu G}}

\begin{document}
%\begin{CJK}{UTF8}{gbsn}

\title{Energy Dependence of Synchrotron X-Ray Rims in Tycho's Supernova
Remnant}

\author{
Aaron Tran \altaffilmark{1,4},
% 陳 doesn't display?
% Also breaks emulateapj bibliography... toggle CJK env w/ AASTeX only
%Aaron Tran (陈宏裕)\altaffilmark{1,4},
Brian J. Williams\altaffilmark{1,5},
Robert Petre\altaffilmark{1},
Sean M. Ressler\altaffilmark{2},
Stephen P. Reynolds\altaffilmark{3}
}

\affil{
\tsup{1}NASA Goddard Space Flight Center, Greenbelt, MD 20771, USA \\
\tsup{2}Dept. Physics, University of California, Berkeley, CA 94720, USA \\
\tsup{3}Dept. Physics, North Carolina State University, Raleigh, NC 27695, USA
}

\altaffiltext{4}{CRESST, University of Maryland, College Park, MD 20742}
\altaffiltext{5}{NASA Postdoctoral Program Fellow}

\begin{abstract}
Several young supernova remnants exhibit thin X-ray bright rims of synchrotron
radiation at their forward shocks.  Thin rims have been taken to indicate that
shock-accelerated electrons rapidly cool downstream of the shock, requiring
strong magnetic field amplification.  But, magnetic field damping immediately
behind the shock could produce similarly thin rims.  We measured rim widths
around Tycho's supernova remnant in 5 energy bands using an archival $750
\unit{ks}$ \Chandra observation.  Rims consistently narrow with increasing
energy and are well-described by either loss-limited or magnetically damped
scenarios, hence X-ray rim widths alone cannot discriminate between models.
Loss-limited fits require amplified magnetic fields of $\abt200$--$1000 \muG$
and particle diffusion coefficients $\abt1$--$100\times$ Bohm values,
consistent with prior work on Tycho's SNR, whereas magnetic damping permits
magnetic fields of $\abt 10$--$300 \muG$ and comparatively smaller diffusion
coefficients (typically $\abt 1$--$10$).  Diffusion coefficients, magnetic
field strengths, and magnetic damping lengthscales are somewhat degenerate in
setting observed rim widths and drop-offs.  Although we cannot discriminate
from width measurements alone, a combined study of X-ray and radio profiles may
break the observed degeneracies to confirm or rule out turbulent magnetic field
damping downstream of supernova remnant shocks.
\end{abstract}

% Six keywords, alphabetical order
\keywords{acceleration of particles ---
    ISM: individual objects (Tycho's SNR) ---
    ISM: magnetic fields ---
    ISM: supernova remnants ---
    shock waves ---
    X-rays: ISM}

% TODO GENERAL WARNING -- DATA ARE NOT ALL FULLY CONSISTENT AT THIS TIME.
% (spectra/profiles/fwhms/fits).

% TODO terminology for "loss-limited" / "damped" / "constant field" / etc
% should be well established

% ============
% Introduction
% ============
\section{Introduction} \label{sec:intro}

% What are these shock rims, what is this acceleration?
Forward shock accelerated electrons in young supernova remnants (SNRs) emit
synchrotron radiation strongly in the shock's immediate wake but quickly turn
off downstream, producing a shell-like morphology with bright X-ray and radio
rims/filaments due to line-of-sight projection \citep{koyama1995}.  The
relatively strong synchrotron radiation, in conjunction with multiwavelength
spectra (radio, X-ray, gamma ray) \citep{aharonian2004, acero2010,
ackermann2013} and observations of rapid time variability \citep{uchiyama2007},
suggests that electrons are accelerated to PeV energies in young SNRs.
Although synchrotron emission due to accelerated electrons does not imply
acceleration of an unseen hadronic component, the prevailing theory of
diffusive shock acceleration (DSA) should operate on both protons and
electrons.  Efficient hadron acceleration in supernova remnant shocks is a
prime candidate source for galactic cosmic rays up to the cosmic ray spectrum's
``knee'' \citep{vink2012}.

% What can we learn from these shocks?  Why are they interesting?
Spectral and spatial measurements of synchrotron rims can constrain physical
parameters of the shock acceleration process.  Many previous workers have
inferred strong magnetic field amplification at SNR shocks, as electrons must
rapidly radiate energy to account for the thinness of observed rims
\citep{bamba2003, vink2003, parizot2006}.  Furthermore, broadband synchrotron
synchrotron flux from supernova remnants steepens at X-ray energies
\citep{reynolds1999}, corresponding to a cut-off in the underlying electron
spectrum as predicted by diffusive shock acceleration \citep{webb1984}.  Many
fundamental questions about the shock acceleration process remain to be
answered.  Under what conditions do shocks accelerate particles efficiently?
How are magnetic fields amplified in such shocks?  \citet{reynolds2008} reviews
relevant observations and open questions to date.  The answers to these
questions are broadly relevant to many astrophysical shocks, such as Earth's
bow shock, stellar wind-blown bubbles, starburst galaxies \citep{heckman1990},
jets of active galactic nuclei, galaxy clusters \citep{van-weeren2010}, and
even cosmological shocks \citep{ryu2008}.
% Expand on these questions -- plasma turbulence, Alfven waves, wave damping
% behind the shock, upstream/downstream plasma interactions and instabilities
% These in turn affect plasma, CR dynamics -- how does the shock propagate and
% interact with the ISM?  Will we see a precursor due to diffusion and/or CR
% streaming?  How efficiently are CRs accelerated?

% What makes these shock rims?
The widths of these X-ray rims are controlled by a combination of
synchrotron losses, particle transport, and magnetic fields immediately
downstream of the shock.  Synchrotron losses depend on the initial electron
energy distribution and the gradual decrease of electron energies downstream of
the shock.  High energy electrons near the shock efficiently radiate harder
synchrotron X-ray photons, but decline in energy as they are transported away
from the shock; this may be seen in spectral softening downstream of the shock
\citep[e.g.,][]{cassam-chenai2007}.  The transport process is driven by bulk
plasma advection downstream of the shock and particle diffusion with respect to
bulk advection.  In general, diffusion allows higher energy electrons to
diffuse further upstream or downstream than would be expected from pure
advection.  Finally, the magnetic field strength may be quickly damped
downstream of the shock, so that electrons may not radiate as efficiently and
thin rims reflect the spatial structure of the field rather than efficient
particle acceleration (and losses) and/or efficient synchrotron cooling
\citep{pohl2005}.

% Whatever makes these shock rims also imposes energy dependence
Moreover, these controlling mechanisms predict different scalings for rim width
as a function of energy.  In the limiting case of constant magnetic field and
negligible particle diffusion, the rims are said to be synchrotron loss-limited
and rim widths should narrow with increasing photon frequency $\nu$ as
$\nu^{-1/2}$. Diffusion smears out rims at all energies with stronger effect at
higher energies, yielding a slower drop-off in rim widths when diffusion
dominates.  Magnetic damping, if at a length scale comparable to filament
widths, also favors more energy-insensitive rim widths -- intuitively, if the
magnetic field turns off, synchrotron radiation turns off regardless of
electron energy.

% Magnetic damping?
Magnetic damping immediately behind the shock offers a tantalizing alternative
explanation for filament rims that requires less extreme magnetic field
amplification, and the possibility of damping in SNR shocks has not been fully
tested \citep{pohl2005, marcowith2010}.  \citet{cassam-chenai2007} examined
radio and X-ray synchrotron rim profiles, but found that neither damping nor
synchrotron losses matched radio profiles particularly well.  \citet{araya2010}
measured rims at multiple energy bands in Cas A, but suggested that the rims
only narrow more weakly than expected and did not consider magnetic damping.
\citet{rettig2012} gave model predictions for several historical SNRs and
proposed discriminating based on filament spectra -- the expectation is that
damped spectra are softer, loss-limited harder.  Recently, \citet{ressler2014}
further extended the idea of discriminating between damped or loss-limited
rims by measuring rim width energy dependence in X-ray energies and found that
rims consistently narrowed in the remnant of SN 1006, favoring synchrotron
loss-limited rims.

% Why Tycho's SNR
To further test these models, we follow \citet{ressler2014} (hereafter,
\citetalias{ressler2014}) by measuring rim widths at multiple energies in the
remnant of Tycho's 1572 supernova (hereafter, Tycho).  Tycho exhibits an
extensive shell of synchrotron-dominated thin rims around its periphery; the
rims show very little thermal emission, consistent with expansion into a low
density ISM \citet{williams2013}.  Tycho, being a nearby galactic SNR, is close
enough that its rims are easily resolved by \textit{Chandra} spatial
resolution. A deep $750 \unit{ks}$ exposure of the entire remnant from 2009
(PI: J. Hughes) also allows us to sample the remnant rims quite finely.
Furthermore, Tycho has been a prime target of study for determining whether
hadrons are being accelerated in historic SNRs \citep[and references
therein]{morlino2012}; understanding the acceleration process will also bear
upon our understanding of CR acceleration in Tycho.

% Paper roadmap
Our procedure closely follows that of \citetalias{ressler2014}.
We selected and measured rim widths at several locations around Tycho's forward
shock, verifying that selected rims are free of thermal line emission.  Using a
model for particle transport with an exponentially cut-off injected electron
spectrum, we fit width measurements to model predictions in the loss-limited
and magnetic damping scenarios to obtain estimates on magnetic field strength
and diffusion coefficient magnitude.  We discuss degeneracy in our model fits
and the implications for future studies of young supernova remnant shocks.

% =============================
% Transport models, observables
% =============================
\section{Nonthermal rim modeling}\label{sec:models}

% Could just discuss one model...
%We consider two particle transport models that we fit to measurements of
%Tycho's thin rim widths.  We measure the rim full widths at half maximum
%(FWHMs) (Section~\ref{sec:fwhms}) and compare our measurements to the transport
%model predictions of rim FWHMs at multiple photon energies, for various input
%parameters.  The models accept four free parameters: magnetic field strength
%$B_0$, a scaled diffusion coefficient $\eta_2$, a diffusion-energy scaling
%exponent $\mu$, and a magnetic damping lengthscale $a_b$ (an $e$-folding
%length).

% Twiddle-ables in our model (the most useful part, since just modeling
% alone don't do much)
Our model fitting explores the space of varying diffusion (a bulk
representative of the underlying MHD turbulence and what have you)
and considers the two scenarios of constant magnetic field (resulting in
so-called loss-limited rims) and a damped field that falls off exponentially
behind the forward shock, as suggested by \citet{pohl2005}.
% this might go at the front before we launch into model details.

\subsection{Particle transport}\label{sec:transport}

% Transport fundamentals
The energy and space distribution of electrons at a supernova remnant's forward
shock controls the synchrotron rims we see in X-ray and radio.  We assume that
diffusive shock acceleration creates a power-law distribution of electrons with
an exponential cut-off at the forward shock, located at $x = 0$ with $x > 0$
increasing downstream of the shock, given as
$f(E,x=0) = K_0 E^{-s} e^{-E/\Ecut}$.  Here $K_0$ is an arbitrary
normalization, $E$ is electron energy, and the cut-off energy $\Ecut$ is given
below in Section~\ref{sec:diffcoeff}; the spectral index $s$ is taken from
the radio index $\alpha$ as $s = 2\alpha + 1$.  \citet{zirakashvili2007} have
derived from DSA a slightly different energy spectrum with super-exponential
cut-off $e^{-(p/p_{\mt{cut}})^2}$, which may impact our results; we use a
simple exponential cut-off for simplicity and consistency with
\citetalias{ressler2014}.

All constants and equations are given in CGS (Gaussian) units.  Our
presentation is somewhat abbreviated, but a fuller exposition and literature
review is given by \citetalias{ressler2014}.

We model electron advection and diffusion downstream of the shock (i.e., for
$x>0$ only) with a 1-D steady-state plane transport equation:
\begin{equation} \label{eq:model}
    v_d \frac{\ptl f}{\ptl x}
    - \frac{\ptl}{\ptl x} \left(D\frac{\ptl f}{\ptl x}\right)
    - \frac{\ptl}{\ptl E} \left(bB^2E^2f\right)
    = K_0 E^{-s} e^{-E/\Ecut} \delta(x) ;
\end{equation}
this follows the work of \citet{webb1984, berezhko2004, cassam-chenai2007,
morlino2010, rettig2012}.  Here $f = f(E,x)$ is electron distribution as a
function of electron energy $E$ and downstream distance $x$, $D$ is the
diffusion coefficient, and $v_d$ is plasma velocity downstream of the shock.
The constant $b = 1.57 \times 10^{-3}$ in CGS units arises from synchrotron
power loss at electron energy $E$ (i.e., $\ptl E/\ptl t = b B^2 E^2$).
We have assumed isotropic diffusion and consider only processes operating
downstream of the shock.  More sophisticated treatments may explicitly treat
magnetic fields and diffusion across the shock, plasma streaming instabilities
and particle losses, anisotropic diffusion, the DSA process and efficiency, and
other details in great depth \citep[e.g.,][and references therein]{reville2013,
bykov2014, ferrand2014}.

To determine rim profiles and widths, we numerically solve for electron
distribution $f(E,x)$ using Green's function solutions by \citet{lerche1980}
and \citet{rettig2012}; the solutions are given by \citetalias{ressler2014}
using notation similar to ours.  We note, however, that the solutions for
$f(E,x)$ in equation~\eqref{eq:model} are derived assuming that $D(x) B^2(x)$
is constant \citep{lerche1980}; we will revisit this assumption further below.
% TODO give section number when you revisit this
The electron distribution may be integrated over the one-particle synchrotron
emissivity $G(y)$ to obtain the ``total'' emissivity:
\begin{equation} \label{eq:emissivity}
    j_{\nu}(x) \propto \int_0^\infty G(y) f(E,x) dE
\end{equation}
where $y \equiv \nu/(c_1 E^2 B)$ is a scaled synchrotron frequency and
$G(y) = y \int_y^\infty K_{5/3}(z) dz$ with $K_{5/3}(z)$ a modified Bessel
function of the second kind \citep{pacholczyk1970}.  Integrating emissivity
over lines of sight for a spherical remnant yields intensity as a function of
radial coordinate $r$:
\begin{equation} \label{eq:intensity}
    I_{\nu}(r) = 2 \int_0^{\sqrt{r_s^2 - r}}
                   j_{\nu} \left( r_s - \sqrt{s^2 + r^2} \right) ds
\end{equation}
where $s$ is the line-of-sight coordinate.  Finally, we compute the full width
at half maximum (FWHM) of the resulting intensity profile to compare to our
measurements of rim width.

% Diffusion
\subsection{Diffusion energy-dependence} \label{sec:diffcoeff}

% Bohm, Bohm-like, and MHD/turbulence-inspired diffusion
Most previous work has assumed Bohm-like diffusion in plasma downstream of SNR
shocks.  Bohm diffusion assumes that the particle mean free path $\lambda$ is
equal to the gyroradius $r_g = E/(eB)$, yielding diffusion coefficient
$D_{\mt{B}} = \lambda c / 3 = c E / (3 e B)$; here $c$ is the speed of light,
$E$ is particle energy, $e$ is the elementary charge, and $B$ is magnetic
field.  Bohm-like diffusion encapsulates diffusion scalings of $D \propto E$,
introducing a free prefactor $\eta$ such that $\lambda = \eta r_g$ allows for
varying diffusion strength.  However, Bohm diffusion at $\eta = 1$ is commonly
considered a lower limit on the diffusion coefficient at all energies.

Different types of MHD turbulence may give rise to a non-linear energy
scaling for diffusion coefficient instead \citep{giacalone1999, reynolds2004}.
Power-wavenumber spectra for Kolmogorov and Kraichnan turbulence, with a
quasi-linear approximation for particle motion,
\note[Aaron]{I don't really know what this means}
predict $D \propto E^{1/3}$ and $D \propto E^{1/2}$ respectively.
Particle gyroradii larger than field coherence length may also lead to
$D \propto E^{2}$ \note[Aaron]{Source?? Parizot mentions this in one sentence
and cites nothing.  I have no idea what this means.}

% Define our diffusion coefficient, introduce fiducial energy
In our models, we consider a generalized diffusion coefficient with arbitrary
power-law dependence upon energy following, e.g., \citet{parizot2006}:
\begin{equation} \label{eq:diffcoeff}
    D(E) = \frac{\eta C_d E^\mu}{B}
         = \eta_h D_{\mt{B}}\left(E_h\right) \left(\frac{E}{E_h}\right)^\mu
\end{equation}
where $\mu$ parameterizes diffusion-energy scaling and $\eta$ now has units
of $\mt{erg}^{1-\mu}$.  The right-hand side of equation~\eqref{eq:diffcoeff}
introduces $\eta_h$, a dimensionless diffusion coefficient scaled to the Bohm
value at a fiducial particle energy $E_h$.  Note that $\eta_h$ and $\eta$ are
related as $\eta = \eta_h (E_h)^{1-\mu}$, and $\eta = \eta_h$ for Bohm-like
diffusion ($\mu = 1$).

% Set choice of fiducial energy, eta_2
For subsequent analysis and model fitting, we take fiducial electron energy
$E_2 = E_h$ corresponding to a $2 \unit{keV}$ synchrotron photon and report all
model fit results in terms of $\eta_2 = \eta_h$.  Although the value of $E_2$
varies with magnetic field $B$ as $E_2 \propto B^{-1/2}$ and may vary within
the remnant, tying $\eta_h$ to a fixed observation energy gives a convenient
and intuitive sense of diffusion strength regardless of the underlying electron
energies.  We also note that $\eta = \eta_2 (E_2)^{1-\mu}$.

% e- cutoff energy result
With a diffusion coefficient in hand, we may determine the electron spectrum
cut-off energy by equating synchrotron loss and diffusive acceleration
timescales.  For low energies and small synchrotron losses (cooling time longer
than acceleration time), electrons may be efficiently accelerated; near or
above the cut-off energy, electrons will radiate or escape too rapidly to be
accelerated to higher energies and the energy spectrum drops off steeply.  The
cut-off energy is given as:
\begin{align} \label{eq:ecut}
    \Ecut =
        &\left(8.3\unit{TeV}\right)^{2/(1+\mu)}
        \left(\frac{B_0}{100 \muG}\right)^{-1/(1+\mu)} \nonumber \\
        &\times \left(\frac{v_s}{10^8 \unit{cm\;s^{-1}}}\right)^{2/(1+\mu)}
        \eta^{-1 / (1+\mu)} .
\end{align}
This result is derived by \citet{parizot2006} for $\mu=1$ assuming a strong
shock with compression ratio 4 and isotropic magnetic turbulence on both sides
of the shock.

% srcut constraint on diffusion
As the DSA imposed electron cut-off results in a roll-off on SNR synchrotron
flux, the synchrotron roll-off frequency may provide an independent observable
to estimate shock diffusion and is given by:
\begin{align} \label{eq:rolloff}
    \nu_{\mt{cut}} =
        &c_m \left(13.3 \unit{erg}\right)^{\frac{4}{1+\mu}}
        \left(0.2657 \unit{erg^2\;G}\right)^{-\frac{1-\mu}{1+\mu}} \nonumber \\
        &\left( 100 \muG \right)^{\frac{1-\mu}{1+\mu}}
        \left( \frac{v_s}{10^8 \unit{cm/s}} \right)^{\frac{4}{1+\mu}}
        \left( \eta_2 \right)^{-\frac{2}{1+\mu}} .
\end{align}
In particular, this equation is independent of magnetic field $B$ for all
values of $\mu$ (but, recall that the specific electron energies associated
with $\eta_2$ will depend on $B$).

% Magnetic damping
\subsection{Magnetic fields and damping}

We consider two scenarios for post-shock magnetic field: (1) a constant field
$B(x) = B_0$, corresponding to the loss-limited rim scenario, and (2) an
exponentially damped field of form:
\begin{equation} \label{eq:bdamp}
    B(x) = \left(B_0 - \Bmin\right) \exp\left(-x / a_b\right)
           + \Bmin
\end{equation}
where $a_b$ is an $e$-folding damping lengthscale, following \citet{pohl2005}.
A typical lengthscale for $a_b$ is $10^{16}$ to $10^{17} \unit{cm}$
\citep{pohl2005}, corresponding to $\abt 0.1$--$1\%$ of Tycho's radius.
Hereafter, we report $a_b$ in units of shock radius $r_s$.

The electron cut-off energy (equation~\eqref{eq:ecut}) depends on the magnetic
field, which varies in a damped model.  For simplicity, we assume that the
shock acceleration process is set by the diffusion at the immediate shock and
so equation~\eqref{eq:ecut} stands as evaluated with magnetic field strength
$B_0$.
\note[Aaron]{Is this correct?  Can rederive the expression (particles
penetrate some distance downstream, then back), but I'm not sure if it's even
sensible with diffusion coefficient varying so sharply.}

The solutions to equation~\eqref{eq:model} given by \citet{lerche1980,
rettig2012} assume $D(x) B^2(x)$ constant in order to make the equation
semi-analytically tractable.  Although unphysical, we enforce this criterion by
using a modified diffusion coefficient in our calculations following
\citet{rettig2012}, given as:
\begin{equation} \label{eq:ddamp}
    D(E,x) = \frac{\eta C_d E^\mu}{B_0}
             \left[ \frac{\Bmin}{B_0} +
                    \frac{B_0 - \Bmin}{B_0} e^{-x/a_b} \right]^{-2}
\end{equation}
This has the unfortunate effect of strengthening the spatial-dependence of the
diffusion coefficient, which may or may not be physical.
\note[Aaron]{I am not certain whether this should be scaled by $\Bmin$ instead,
will check shortly}

\subsection{Rim width-energy dependence} \label{sec:energydep}

% Define \mE
Following \citetalias{ressler2014}, we parameterize rim width-energy dependence
in terms of a scaling exponent $\mE$ defined as:
\begin{equation}
    w(\nu) \propto \nu^{\mE}
\end{equation}
where $w(\nu)$ is filament FWHM as a function of observed photon frequency
$\nu$.  The exponent $\mE = \mE(\nu)$ is generally energy dependent.  We may
refer to observed photons interchangeably by energy or frequency $\nu$, but $E$
is reserved for electron energy.

% Intuition: transport lengthscales in the absence of damping
We may consider advective and diffusive lengthscales for electron transport,
which naturally give rise to scaling relations:
\begin{equation} \label{eq:lad}
    l_{\mt{ad}} = v_d \tsynch \propto v_d B_0^{-3/2} \nu^{-1/2}
\end{equation}
and
\begin{equation} \label{eq:ldiff}
    l_{\mt{diff}} = \sqrt{D \tsynch} \propto \eta^{1/2} B_0^{-3/2} \nu^{(\mu-1)/4}
\end{equation}
The characteristic time is the synchrotron cooling time
$\tsynch = 1 / (b B^2 E)$ with $b = 1.57 \times 10^{-3}$.
If both diffusion and magnetic field damping are negligible (whether at a small
range of, or all, electron energies) and electrons are only loss-limited as
they advect downstream, $\mE$ attains a minimum value $\mE = -1/2$; this is the
strongest energy-dependence possible in our models.  Diffusion increases $\mE$
monotonically towards a value between $-1/4$ and $1/4$, depending on the value
of $\mu$.  In practice, $\mE$ appears to plateau at values between $-0.4$ and
$0$, with larger $\mu$ values yielding $\mE$ closer to $0$ (weaker
energy-dependence) \citepalias[Figure 3]{ressler2014}.  This may be attributed
in part to the electron energy cut-off, which limits the available electrons at
higher energies and thus thins the rims slightly, decreasing $\mE$
\citepalias[Figure 5]{ressler2014}.

% Intuition: how damping affects rim widths
We expect magnetic damping to produce comparatively energy-independent rim
widths.  In particular, if synchrotron rims widths are set by magnetic damping
at some observation energy, then the rims must be damped at all lower
observation energies as well.  Then we may expect relatively stable rim widths
below a threshold energy, then decreasing rim widths once advective or
diffusive lengthscales fall below the damping lengthscale at higher photon
energies.

We shall see below that this is qualitatively correct in setting model
predictions of rim shapes (within $\abt a_b$ of the shock front), but our
results are complicated by the increasing importance of energy-dependent
synchrotron losses downstream of the thin rim.  As we use FWHM as a singular
proxy for rim shapes, our procedure is not able to discern the transition in
shape downstream of the damped rim that may impact profile FWHMs, even as
damping controls the main observed rim emission.

% A comment
It may be worth noting that this rim width-energy dependence is a morphological
manifestation of the spectral softening downstream of the forward shock, as
observed by, e.g., \citet{cassam-chenai2007}.  So we are simply taking somewhat
of a different approach to characterizing this softening.

% ============
% Observations
% ============
\section{Observations}
\label{sec:observations}

% NOTE go through and double check tenses.
We measured synchrotron rim full widths at half maximum (FWHMs) from an
archival \Chandra ACIS-I observation of Tycho
(RA: 00\tsup{h}25\tsup{m}19\fs0, dec: +64\arcdeg08\arcmin10\farcs0; J2000)
between 2009 Apr 11 and 2009 May 5 (PI: J. Hughes;
\dataset[ADS/Sa.CXO\#obs/10093--10097]{ObsIDs: 10093--10097},
\dataset[ADS/Sa.CXO\#obs/10902--10906]{10902--10906}).
The total exposure time was $734 \unit{ks}$.
Level 1 \Chandra data were reprocessed with CIAO 4.6 and CALDB 4.6.1.1 and kept
unbinned with ACIS spatial resolution $0.492\arcsec$.
Merged and corrected events were divided into five energy bands:
0.7--1 keV, 1--1.7 keV, 2--3 keV, 3--4.5 keV, and 4.5--7 keV.
We excluded the 1.7--2 keV energy range to avoid any \ion{Si}{13} (He$\alpha$)
emission at 1.85 keV, prevalent in the remnant's thermal ejecta, which
might contaminate our nonthermal profile measurements.

% NOTE keep region/filament numbering up to date
We selected 20 regions in 5 distinct filaments around Tycho's shock
(Figure~\ref{fig:snr}) based on the following criteria: (1) filaments should be
clear of spatial plumes of thermal ejecta in \Chandra images, which rules out,
e.g., areas of strong thermal emission on Tycho's eastern limb; (2) filaments
should be singular and localized; multiple filaments should either not
overlap or completely overlap (rules out parts of NE limb); (3) filament
peaks should be evident above the background signal or downstream thermal
emission (rules out faint southern filaments).  We accepted several
regions with poor quality peaks in the lowest energy band (0.7--1 keV) so long
as peaks in all higher energy bands were clear and well-fit.
\note[Aaron]{vague, imprecise}
We grouped regions into filaments by visual inspection of the remnant.
Within each filament, we chose region widths to obtain comparable counts
at the thin rim peak.
All measured rim widths are at least $1\arcsec$ and hence are resolved by
\textit{Chandra}'s point-spread function (PSF), which has FWHM of
$\abt 1\arcsec$ at $4\arcmin$ off-axis.

% TODO final figure should be higher quality, use vector overlay if possible
\begin{figure}
    \centering
    \iftoggle{manuscript}{
        \includegraphics[width=0.5\textwidth]{figures/f0-snr-inv.png}
    }{
        \plotone{figures/f0-snr-inv.png}
    }
    \caption{RGB image of Tycho with region selections overlaid.  Image bands
    are 0.7--1 keV (red), 1--2 keV (green) and 2--7 keV (blue).  Bold region
    labels (1, 16) indicate region selections shown in
    Figures~\ref{fig:spec},~\ref{fig:profiles}.  Filament 1: Regions 1--3,
    filament 2: regions 4--10, filament 3: regions 11-13, filament 4:
    regions 14--17, filament 5, regions 18--20.
    \label{fig:snr}}
\end{figure}
% TODO TODO TODO UPDATE REGION # BOLDING TO MATCH DISPLAYED FILAMENTS ETC

% ----------------
% Filament spectra
% ----------------
\subsection{Filament spectra}
\label{sec:spec}

% NOTE be consistent in usage of word 'sections'
We extracted spectra for all regions to check that rim width measurements are
not affected by contaminating thermal line emission.  In each region, we
selected upstream and downstream sections from which to extract spectra.  The
upstream section is the smallest sub-region that contains the measured FWHM
bounds from all energy bands (see Section~\ref{sec:fwhms}); i.e., this section
overlies the thin rim.  The downstream section extends from the back of the rim
(end of the upstream section) to the rim model fit's (equation~\eqref{eq:prof})
downstream limit.  We extracted spectra from all \Chandra ObsIDs and grouped
the data into bins of $\geq 15$ counts.  Figure~\ref{fig:spec} plots an example
profile ($4.54$--$7 \unit{keV}$) with the downstream and upstream sections
highlighted.  We extracted background spectra from circular regions (radius
$\abt 30\arcsec$) around the remnant's exterior; each thin rim region's spectra
subtracted the closest background region's spectrum.
% Background region size is valid for data-tycho/bkg-2/ selections

% NOTE must match fig:snr, fig:profiles!
% Keep region numbers updated and consistent.
\begin{figure*}
    \iftoggle{manuscript}{
        \epsscale{0.7}
    }{}
    \plotone{figures/spec_01.pdf} \\
    \plotone{figures/spec_16.pdf}
    \caption{Spectra and fits from Regions 1 (top) and 16 (bottom) show varying
    rim morphology and example of rim where 0.7--1 keV peak could not be fit.
    Left: $4.5$--$7 \unit{keV}$ profile with highlighted downstream (blue) and
    upstream (grey) sections.  Intensity in arbitrary units (a.u.).  Middle:
    downstream spectrum with absorbed power law fit; Si and S lines at $1.85$,
    $2.45 \unit{keV}$ are clearly visible.  Right: upstream spectrum with
    absorbed power law fit shows that each region is likely free of thermal
    line
emission.}
    \label{fig:spec}
\end{figure*}

% Upstream spectra and numbers
% NOTE keep stated ranges of spectral indices, column densities up to date
We fit spectra from each region's upstream and downstream sections to an
absorbed power law model (XSPEC 12.8.1, \texttt{phabs*po}) between $0.5$--$7
\unit{keV}$ with photon index $\Gamma$, hydrogen column density $N_{\mt{H}}$,
and a normalization as free parameters.  Table~\ref{tab:spec} lists best fit
parameters and reduced $\chi^2$ values for all regions.  Spectra from thin rims
(upstream sections) are well-fit by the power law model alone; the best-fit
photon indices ($2.4$--$3$) and column densities ($0.6$--$0.8 \times 10^{22}
\unit{cm^{-2}}$) are consistent with previous spectral fits to Tycho's
nonthermal rims \citep{hwang2002, cassam-chenai2007}.

% Downstream spectra and numbers
% NOTE keep stated ranges, region numbers up to date
% THIS IGNORES THE REGION 19 SPECTRUM FIT (!!)
Spectra from sections downstream of the rims are generally poorly-fit,
reflecting thermal contamination from \ion{Si}{13} and \ion{S}{15} He$\alpha$
line emission at $1.85$ and $2.45$ keV.  To confirm that thermal emission is
dominated by these two lines near the shock, we also performed fits to
downstream spectra with both lines excised ($1.7$--$2.0 \unit{keV}$,
$2.3$--$2.6 \unit{keV}$ removed) or with both lines fitted to Gaussian
profiles.  When the lines are excised from our fits, we obtain
$\chi^2_{\mt{red}}$ values between $0.8$--$1.4$ excepting only Regions 5 and 20
with $\chi^2_{\mt{red}} \sim 1.8$ due to the presence of many additional lines.
When lines are fitted to Gaussian profiles, we find $\chi^2_{\mt{red}}$ values
$0.83$--$1.22$, again excepting Regions 5, 20 which have $\chi^2_{\mt{red}}
\sim 1.6$.  In both fits (lines excised or modeled), we find somewhat smaller
best fit column densities ($0.32$--$0.74 \times 10^{22} \unit{cm^{-1}}$),
but best fit photon indices ($2.52$--$3.05$) are consistent with those of the
upstream spectra, confirming that the same synchrotron continuum may be
recovered from beneath thermal line emission.

\begin{table}
    \scriptsize
    \centering
    \caption{Region spectra fit parameters\label{tab:spec}}
    \input{tab-spec.tex}
\end{table}

% This table is still useful, just to look at the parameters
% (or to send around to collaborators, if needed)
%\begin{table*}
%    \footnotesize
%    \centering
%    \caption{Region spectra fit parameters\label{tab:spec-pt2}}
%    \input{tab-spec-pt2.tex}
%\end{table*}

Our spectral fitting confirms that all selected region are practically free of
thermal line emission, as already suggested by visual inspection
(Figure~\ref{fig:snr}).  Our exclusion of all $1.7$--$2 \unit{keV}$ photons
further limits thermal contamination as $1.85 \unit{keV}$ Si line emission is
over a third of Tycho's thermal flux as detected by \Chandra \citep{hwang2002}.

% --------------------------
% FWHM measurement procedure
% --------------------------
\subsection{Filament width measurements}
\label{sec:fwhms}

% NOTE keep mentioned number of regions updated!
% Check these (arcsec) numbers before final version.
We obtained radial intensity profiles from $\abt 10$--$20\arcsec$ behind the
shock to $\abt 5$--$10\arcsec$ in front for each of our regions in all five
energy bands.  To increase signal-to-noise, we integrate along the shock
($5$--$23\arcsec$) in each region.  Plotted and fitted profiles are reported in
vignetting and exposure-corrected intensity units; error bars are computed from
raw counts assuming Poisson statistics.  Intensity profiles peak sharply
within $\abt 2$--$3\arcsec$ behind the shock, demarcating our thin rims,
then fall off gradually until thermal emission picks up further behind the
shock.

We fitted rim profiles to a piecewise two-exponential model:
\begin{equation} \label{eq:prof}
    h(x) =
    \begin{cases}
        A_u \exp \left(\frac{x_0 - x}{w_u}\right) + C_u, &x \geq x_0 \\
        A_d \exp \left(\frac{x - x_0}{w_d}\right) + C_d, &x < x_0
    \end{cases}
\end{equation}
where $h(x)$ is profile height and $x$ is radial distance from remnant center.
The rim model, which is strictly empirical, has 6 free parameters:
$A_u, x_0, w_u, w_d, C_u$, and $C_d$; $A_d = A_u + (C_u - C_d)$ enforces
continuity at $x=x_0$. Our model is similar to that of \citet{bamba2003,
bamba2005-hist} and differs slightly from that of \citetalias{ressler2014}.
To fit only the nonthermal rim in each intensity profile, we selected the fit
domain for each profile as follows.  The downstream bound was set at the first
local data minimum downstream of the rim peak, identified by smoothing the
profiles with a 21-point ($\abt 10\arcsec$) Hanning window.  The upstream bound
was set at the profile's outer edge (i.e., no data were removed).

\begin{figure*}%[ht]
    \iftoggle{manuscript}{
        \epsscale{1}
    }{}
    \plotone{figures/prfs_01.pdf}
    \plotone{figures/prfs_16.pdf}
    \caption{Best fit profiles with measured FWHMs demarcated for each energy
        band in Region 1 (top) and Region 16 (bottom).  Energy bands increase
        from left to right.  Data points in red were excluded from profile
        fitting domains as described in text.}
    \label{fig:profiles}
\end{figure*}

% NOTE keep blacklisted/excluded region count updated!
From the fitted profiles we extracted a full width at half maximum (FWHM) for
each region and each energy band after subtracting a constant background term
$\min(C_u, C_d)$.  We could not resolve a FWHM in 4 of 20 regions at 0.7--1 keV
(Table~\ref{tab:fwhms}); in these regions, either the downstream FWHM bound
would extend outside the fit domain or we could not find an acceptable fit to
equation~\eqref{eq:prof}.  We were able to resolve FWHMs for all regions at
higher energy bands (1--7 keV).
% NOTE review the FWHM rejection criterion, on the final set of regions
% We will be blacklisting by hand, and we have to explain how we chose the
% blacklisted ones.

To estimate FWHM uncertainties, we horizontally stretched each best-fit
profile by mapping radial coordinate $x$ to
$x'(x) = x (1 + \xi (x-x_0)/(50\arcsec-x_0))$ with $\xi$ an arbitrary stretching
parameter and $x_0$ the best-fit rim center from equation~\eqref{eq:prof};
this yields a new profile $h'(x) = h(x'(x))$.
We varied $\xi$ (and hence rim FWHM) to vary each profile fit $\chi^2$ by 2.7
and took the stretched or compressed FWHMs as upper/lower bounds on our
reported FWHMs.

% -------------
% Model fitting
% -------------
\subsection{Filament model fitting}
\label{sec:fits}

% How we ran model fits (itemize knobs, explain energy band mapping, errors,
% fitting procedure
We fit model FWHM predictions given by Equation~ \eqref{eq:model} to our
measured rim widths as a function of energy by varying several physical
parameters: magnetic field strength $B_0$, normalized diffusion coefficient
$\eta_2$, diffusion-energy scaling exponent $\mu$, and minimum field strength
$\Bmin$ and lengthscale $a_b$ for a damped magnetic field scenario.  We
mapped each width measurement to the lower energy limit of its energy band;
e.g., $0.7$--$1 \unit{keV}$ is assigned to $0.7 \unit{keV}$ and fitted to model
profile widths at $0.7 \unit{keV}$.  Errors in our least squares fits average
the positive and negative errors on each FWHM measurement.  For an input set of
model parameters, we numerically computed intensity profiles and hence model
FWHMs as detailed in Section~\ref{sec:transport}, and used a
Levenberg-Marquardt fitting routine to identify best fit model parameters
(subject to varied constraints on model parameters in the following paragraph).
To assist the nonlinear fitting, we tabulated model FWHM values on a large grid
of model parameters (primarily $\eta_2$ and $B_0$) and used best-fit grid as
initial guesses for fitting.
% TODO this may be obsoleted if we recompute stuff...  I only computed errors
% on the loss-limited model, and those fits are now out of date too...
Errors on best fit parameters were computed by varying each fit parameter and
obtaining a new best fit model, with one less degree of freedom, to find the
parameter limit such that $\Delta \chi^2 = 1$ for a roughly 1-$\sigma$ error.

% How we twiddled model knobs
We first considered a purely loss-limited model (constant magnetic field
$B_0$) with three parameters $\mu$, $\eta_2$, and $B_0$.  To make nonlinear
fitting tractable, we fixed $\mu$ in all fits and considered $\mu = 0$, $1/3$,
$1/2$, $1$, $1.5$, and $2$; recall that $\mu = 1/3$ and $1/2$ correspond to
Kolmogorov and Kraichnan turbulence.  The remaining parameters $\eta_2$ and
$B_0$ varied freely.

% Damped case
For damped magnetic field fits, we held the asymptotic field strength
$\Bmin$ inside the remnant constant at $5 \muG$, slightly higher than
typical intergalactic values of $\abt 2$--$3 \muG$ \citep{lyne1989, han2006}.
We fixed $a_b$ at $12$ different values and ran fits with $\eta_2$ and $B_0$
free.  Formally, we do not seek a best fit value of damping length $a_b$, so
our ``best'' damping fits report the value of $a_b$ yielding the smallest
$\chi^2$ value.  Finally, we fixed $\mu = 1$ for all damped fits.

% Variants on our fitting procedure
To better explore the parameter space of model predictions, we also computed
model fits with both $\eta_2 = 1$ and $\mu = 1$ fixed for both scenarios of
constant and damped magnetic field.  We also attempted damping fits with
$\Bmin = 2 \muG$, somewhat on the lower end of estimated galactic
magnetic field values.

We required $\eta_2$ to be positive in all fits.  We also deemed
best-fit and error bound values with $\eta_2 \geq 10^5$ and $B_0 \geq 10
\unit{mG}$ to be effectively unconstrained. % Change this if needed.

% Tycho parameters
A few remnant-specific parameters affect the model calculations.  We adopted
electron spectral index $s = 2.16$ from radio spectral index $\alpha = 0.58$
\citep{sun2011}, remnant distance $3 \unit{kpc}$ \citep[cf.][]{hayato2010}, and
shock radius $1.08 \times 10^{19} \unit{cm}$ from angular radius $240\arcsec$
\citep{green2014}.  Tycho's forward shock velocity varies with azimuth by up to
a factor of 2; we interpolated velocities reported by \citet{williams2013}
(rescaled to $3 \unit{kpc}$) to estimate individual shock velocities for each
region.

% Model resolution error
As the full model must be numerically solved, its predicted rim widths are
subject to resolution error in the numerical integrals (discretization over
radial coordinate, line-of-sight coordinate, electron distribution, Green's
function integrals).  We chose integration resolutions such that the
fractional error in model FWHMs associated with halving or doubling each integration
resolution (individually) is less than $1\%$ for the parameter space relevant
to our filaments -- $\eta_2 \leq 10^3$ for the loss-limited model and $\eta_2
\leq 10$ for the magnetic damping model).  In our error analysis, the maximum
resolution errors in a sample of parameter space are typically $0.1$--$1\%$,
but mean and median errors are typically an order of magnitude smaller that
maximum errors.  Thus we expect numerical errors to be negligible.

% =======================
% Results, FWHMs and fits
% =======================
\section{Results}

% --------------------
% FWHM results, tables
% --------------------
\subsection{Rim widths}
\label{sec:fwhm-results}

% Present table of FWHM results
Table~\ref{tab:fwhms} reports FWHM measurements for all of our regions.
We also report $\mE$ values for all but the lowest energy band, computed
point-to-point between discrete energy bands as:
\begin{equation}
    \mE(E_2) = \frac{\ln(w_2/w_1)}{\ln(E_2/E_1)}
\end{equation}
where $w_1, w_2$ and $E_1, E_2$ are FWHMs and lower energy values for each
energy band -- e.g., $\mE$ at $1 \unit{keV}$ is computed using FWHMs from
$0.7$--$1 \unit{keV}$ and $1$--$1.7 \unit{keV}$, with $E_2 = 1 \unit{keV}$ and
$E_1 = 0.7 \unit{keV}$.  Errors on $\mE$ are propagated in quadrature from
adjacent FWHM measurements.

\begin{table*}
    \iftoggle{manuscript}{
        \tiny
    }{
        \scriptsize
    }
    \centering
    \caption{Measured full widths at half max (FWHMs) for all regions.
             \label{tab:fwhms}}
    \input{tab-fwhms.tex}
\end{table*}

% Observations of mE and rim widths and stuff
Measured rim widths decrease with energy in most regions and energy bands.
Although the measurement scatter is quite large, shown dramatically in the
point-wise computed $\mE$ values, the mean rim width decreases consistently
with increasing energy.  Furthermore, mean $\mE$ values are consistently
negative and tend smoothly towards $0$ (weaker energy-dependence) with
increasing energy.

% Errors in rim widths
The errors on FWHM measurements are typically $\abt 10\%$ or smaller,
reflecting the high quality of underlying \Chandra data as also seen in
Figure~\ref{fig:profiles}.  Scatter in FWHM measurements may be attributed in
part to (1) our profile fitting procedure, which depends on a somewhat
arbitrary choice of profile fit function, and (2) variation in Tycho's rim
morphology (see Figure~\ref{fig:profiles}).

% -------------------------
% Model fit results, tables
% -------------------------
\subsection{Model fit results}
\label{sec:fit-results}

% Explain our approach
We tabulate and plot detailed results for Regions 1 and 16, the same regions
shown in Figures~\ref{fig:spec} and \ref{fig:profiles}.  Although only a small
subset of our regions, qualitative trends in fit parameters were comparable in
most regions; our most salient observations, discussed below, are well
exemplified by Regions 1 and 16.  Best fit results for all regions are
summarized in Tables~\ref{tab:fits-all}, \ref{tab:fits-all-eta2one}.  In all
tables, we report values of $\chi^2$ rather than $\chi^2_\mt{red}$ to ease
comparison of different fits, as we often manually stepped one or two
parameters that were held constant in fitting.

% Best fit model results and tables for Regions 1, 16.
% Explain a bit about model shapes, and the relative similarity in shapes
Modeled profiles with best fit FWHMs for Regions 1 and 16 are plotted in
Figure~\ref{fig:rims}), and their derived best-fit FWHM-energy curves
are plotted in Figure~\ref{fig:fits} alongside measured FWHMs.
The best fit model profiles do not quite match our measured rim profiles
(Figure~\ref{fig:profiles}).  Our transport equation does not account for
diffusive transport upstream of the shock, projection effects from shock
ripples or other structure, and thermal emission downstream of the thin rim.
Hence, fitting to FWHM measurements should be less sensitive to these effects
as compared to directly fitting entire rim profiles to model predictions.
We also note that although Figure~\ref{fig:rims} plots model profiles for
$\mu = 1$ only, the best-fit FWHM-energy curves show that model profiles for
$\mu \neq 1$ give very similar FWHM-energy dependence.  This is borne out in
our best fit parameters in Tables~\ref{tab:fits-loss} and \ref{tab:fits-damp}.

\begin{figure}
    \centering
    \iftoggle{manuscript}{
        \epsscale{0.5}
    }{}
    \plotone{figures/full-man_err-prfs_01.pdf} \\
    \plotone{figures/full-man_err-prfs_16.pdf}
    \caption{Best fit full model rim profile width predictions for Regions 1
        (top) and 16 (bottom) for loss-limited and damped scenarios with
        $\mu = 1$ fixed; profiles are normalized to intensity $1$.  Shaded
        regions on damped profiles indicate the damping ($e$-folding)
        lengthscale for the magnetic field.  Note that radial coordinate ($r$)
        axes agree in each row (same region), but not in each column.
    \label{fig:rims}}
\end{figure}

\begin{figure}
    \centering
    %\epsscale{0.45}  % For manuscript layout only!
    \iftoggle{manuscript}{
        \includegraphics[width=0.46\textwidth]{figures/full-man_err-energywidth-01.pdf}
        \includegraphics[width=0.46\textwidth]{figures/full-man_err-energywidth-16.pdf}
    }{
        \includegraphics[width=0.23\textwidth]{figures/full-man_err-energywidth-01.pdf}
        \includegraphics[width=0.23\textwidth]{figures/full-man_err-energywidth-16.pdf}
    }
    \caption{Full model width predictions as a function of energy for Regions 1
    (top) and 16 (bottom) with measured data.  Dotted/dashed curves plot
    pure loss-limited fits with varying $\mu$, solid curve plots estimated best
    model fit with moderate damping ($a_b < 0.01$).  Note that ordinate FWHM
    axes do not match and are offset from the origin to better show model
    predictions and data variation.
    \label{fig:fits}}
\end{figure}

% Observations on qualitative trends in fit parameters
% NOTE keep numbers updated.  # of regions "favoring" mu >= 1
% KEEP B0 minimum number updated!!!!!
Loss-limited fits with $\mu \lesssim 1$ generally yield larger parameter values
and errors for both $\eta_2$ and $B_0$, and they are more likely to be
ill-constrained (e.g., $\eta_2 > 10^3$ or $B_0 > 10^3 \muG$).  Fits with $\mu
\gtrsim 1$ yield smaller $\chi^2$ values in a majority of regions ($\abt 17$ of
$20$), although we do not explore the robustness of this observation.  Fits to
the same data with varying $\mu$ can yield $\eta_2$ varying by $1$--$2$ orders
of magnitude; Region 16 illustrates this dramatically
(Table~\ref{tab:fits-loss}).

Magnetically damped rims are able to fit width-energy dependence in our
measurements at least as well as purely loss-limited rims for an acceptable
range of $a_b$ values (Table~\ref{tab:fits-damp}).  Damped model fits show some
variability in best fit parameters, mainly in $\eta_2$, but are generally
better-behaved than their loss-limited counterparts.  Smaller values of $a_b$
permit and favor smaller best fit values of $B_0$ and $\eta_2$ in all regions.

% Example loss-limited fits
\begin{table}
    \scriptsize
    \centering
    \caption{Loss-limited fits for Regions 1, 16
    \label{tab:fits-loss}}
    \input{tab-fits-loss.tex}
    \tablecomments{Model fits computed at several fixed values of $\mu$ with
    $\eta_2$, $B_0$ free and constant magnetic field $B(x) = B_0$.
    Region 1 has 2 degrees of freedom (dofs); Region 16, 3.}
\end{table}

% Example damping fits
\begin{table}
    \scriptsize
    \centering
    \caption{Magnetic damping fits for Regions 1, 16
    \label{tab:fits-damp}}
    \input{tab-fits-damp.tex}
    \tablecomments{Model fits computed at several fixed values of $a_b$ with
    $\eta_2$, $B_0$ free and damped magnetic field (equation~\eqref{eq:bdamp}).
    Fit degrees of freedom are as in Table~\ref{tab:fits-loss}}
\end{table}

% Fit results for ALL regions, mu = 1, loss-limited and damped
% NOTE keep # of dofs updated
\begin{table}
    \tiny
    \centering
    \caption{Best model fits for all regions, $\mu = 1$
    \label{tab:fits-all}}
    \input{tab-fits-all.tex}
    \tablecomments{Loss-limited fits fix $\mu=1$ to match the damping model;
    best damped fits select best $a_b$ value below $0.01$ from the range shown
    in Table~\ref{tab:fits-damp}, enforcing moderate damping.  Fits for Regions
    1, 2, 12, 18 have 2 dofs; all others have 3.  The manual selection of a
    best $a_b$ value may be construed as removing one more dof.}
\end{table}

In Table~\ref{tab:fits-all-eta2one} we also present best loss-limited and
damped fits with $\eta_2 = 1$ fixed.  As before, both loss-limited and damped
models are equally capable of fitting the observed data.

% As above but with eta2=1 fixed!
\begin{table}
    \scriptsize
    \centering
    \caption{Best model fits for all regions, $\mu = \eta_2 = 1$
    \label{tab:fits-all-eta2one}}
    \input{tab-fits-all-eta2one.tex}
    \tablecomments{See comments on Table~\ref{tab:fits-all-eta2one}.  Fits have
    one additional dof as $\eta_2 = 1$ is fixed.}
\end{table}

% ==========
% Discussion
% ==========
\section{Discussion}

Here we first explore trends in our data and best-fit diffusion and magnetic
fields.  We then discuss discrimination of the loss-limited and damping models
(i.e., our ability to favor or constrain a finite value of $a_b$), and possible
paths forward to better constrain our model parameters.

% --------------------------------------------
% Commentary on data, observations, parameters
% Tycho distance estimates, maybe compression ratio stuff...
% Any constraints on parameter values go here
% --------------------------------------------
\subsection{Data and model fit interpretation}

\subsubsection{Width-energy dependence}

% First address the broad trend of leveling energy dependence in mE)
We suggest that our observed decrease in $|\mE|$ with increasing energy is
caused by diffusion, which weakens width-energy dependence at higher energies.
The photon energy at which advection and diffusion are comparably important
($l_{\mt{ad}} = l_{\mt{diff}}$) is exactly the synchrotron roll-off energy,
$\abt 0.3 \unit{keV}$ in Tycho \citep{hwang2002}.  Then we expect diffusion to
begin overtaking advection and depress $\mE$ towards $0$ in soft X-rays, as
opposed to $\mE \abt -0.5$ expected when advection is important
(Section~\ref{sec:energydep}).  A decrease in $|\mE|$ is then reassuringly
consistent with the framework of diffusive shock acceleration.

% This conclusion is based strictly on our measurements.
% Our transport model hard-codes this width leveling through \Ecut

% Discuss variability in width measurements
The large scatter in filament widths and hence point-wise computed $\mE$ values
(Table~\ref{tab:fwhms}) does relegate our interpretation of $\mE$ to conjecture.
Multiple regions show both increases and decreases in filament widths between
energy bands, and we have seen that diffusion ($\eta_2$) is poorly constrained
by our fit results.  Our averaged results should mitigate this natural
measurement spread, however.

% Can damping explain stuff?
The next question is whether damping may also produce this effect, independent
of assumptions on the shock acceleration process.  Conceptually, if damping
controls rim widths past the cut-off energy, $|\mE|$ may slightly decrease or
even increase as rims transition from being damping controlled to being
diffusion loss-limited.  If damping is only important well below the cut-off
energy, then we may see an increase in $|\mE|$ as advection becomes important.
Although magnetic damping is able to fit observed rim widths, our model assumes
diffusive shock acceleration in computing $\Ecut$ and hence we may not use our
model fits to assess whether magnetic damping can independently account for
decreasing $|\mE|$.

% Curious effects of damping competing with other things
Downstream energy losses in a well damped field (multiple $e$-folding
lengths ($a_b$) downstream) may also introduce energy dependence in rim widths.
Sufficiently energetic electrons will radiate and cool within the damping
length $a_b$, but lower energy electrons will perceive a relatively constant
field $\abt \Bmin$ downstream of the shock.  Synchrotron losses will
``sieve'' out higher energy electrons as in the loss-limited model, and
observed rim widths, if larger than $a_b$, may be somewhat loss-limited at low
energies, well-confined by damping at moderate energies, then diffusion (or
cut-off spectrum) limited at the highest energies.

We revisit the effects of damping further below.  The observation of decreasing
$|\mE|$, however, remains consistent with and expected from the theory of
diffusive shock acceleration.

\subsubsection{Modeled diffusion and magnetic field amplification}

% Discuss degeneracy in loss-limited and damped models both
We now turn to our model fit parameters: magnetic field strength $B_0$ and
scaled diffusion coefficient $\eta_2$.  Our loss-limited fit results are poorly
constrained as $B_0$ and $\eta_2$ may covary without strongly altering fit
quality.  If diffusion is the primary control on rim width -- i.e.,
$l_{\mt{diff}} > l_{\mt{ad}}$ and thus rim width $w \sim l_{\mt{diff}}$) --
then $\eta_2$ and $B_0$ become degenerate as the product $\eta^{1/2}
B_0^{-3/2}$ exerts most control on rim width $w$.  If advection is the primary
control on rim width (widths narrow rapidly with energy, at $\mE \sim -0.5$),
then $\eta_2 \ll 1$ becomes unimportant and the fit is better behaved with
effectively one free parameter.  The degeneracy of $B_0$ and $\eta_2$ is
stronger for smaller values of $\mu$, although there appears no clear trend as
to when a particular fit will be ill-constrained.

In the damped model, a similar degeneracy is present but modified somewhat by
the presence of damping.  For small values of $a_b$ (strong damping), increased
diffusion $\eta_2$ can cause rim widths to \emph{narrow}, contrary to
intuition; $B_0$ and $\eta_2$ generally covary in controlling rim widths, but
the relationship is now less clear.  Speculatively, the opposition to
downstream flow may be tied to the spatial variation of the diffusion
coefficient, which gives rise to an additional term $-(\ptl D(x) / \ptl x)
(\ptl f/\ptl x)$ that opposes downstream flow in equation~\eqref{eq:model}.

% NOTE some of this discussion of parameter values could go
% into results instead.

% B0 lower bounds, etc. in loss-limited model
In our loss-limited model fits, the minimum magnetic field strength we observe
throughout the remnant, within stated errors, is $B_0 = 179 \muG$.  The minimum
values of $B_0$ for each filament (1--5) are $179$, $254$, $287$, $458$, and
$243 \muG$.  We observe the smallest $B_0$ values for $\eta_2$ approaching $0$
and $\mu = 2$; intuitively, $\mu = 2$ gives stronger diffusion at higher
energies ($>2\unit{keV}$) where it is most ``needed'' to reproduce the leveling
off of filament widths and so $\mu = 2$ permits a smaller best-fit $\eta_2$ and
hence smaller $B_0$.  Furthermore, when $\eta_2 = 1$ is fixed we find that
$B_0$ is well-constrained in the loss-limited model, falling strictly between
$180$--$500 \muG$ (Table~\ref{tab:fits-all-eta2one}).  In this scenario the
range of $B_0$ values is directly correlated to the observed range of rim
widths.  Our minimum loss-limited values of $B_0$ are consistent with prior
estimates of $\abt 200$--$300 \muG$ for primarily advective transport
\citep{volk2005, parizot2006, morlino2012}.  If loss-limited, Tycho's rims
require strong magnetic field amplification -- at least $\abt 100\times$ the
typical galactic field values of $\abt 2$--$3 \muG$, versus the expected
$4\times$ amplification from a strong shock with compression ratio $r=4$.

The damped model fits permit much smaller values of $B_0$, as expected.
The maximum best-fit $B_0$ value observed is $600 \muG$, with most values
typically below $300 \muG$.  The minimum value of $B_0$ is around $10$--$20
\muG$, which is consistent with shock compression alone.
Fixing $\Bmin = 2 \muG$ instead of $5 \muG$ generally permits smaller
fit $B_0$ values, though both $\eta_2$ and $B_0$ still display a large spread.

% More detail on relevant B0 estimates follows

% \citet{parizot2006} give $200 \muG$ for pure advection and $230 \muG$ for pure
% diffusion; with both effects, they obtain $300$--$530 \muG$ depending on the
% assumed compression ratio and diffusion-energy scaling $\mu$.
% \citet{volk2005} combine a prediction from nonlinear acceleration with actual
% data.  They too suggest something like $300 \muG$.
% \citet{volk2002} give $240 \muG$ from a multiwavelength study.  Too many papers
% from Russian theorists, seriously.
% \citet{morlino2012} do a similar thing with newer data and favor $B_0 > 200
% \muG$ strong magnetic fields -- it also is favored by the hadronic scenario of
% pion decay to help explain gamma-ray observations.
% \citet{cassam-chenai2007} take values $215 \muG$ and $130 \muG$ for different
% particle injection efficiencies to describe the X-ray rims of Tycho; note that
% they find an overall compression ratio $r \approx 6$ and treat the effect of CR
% acceleration and escape on the shock hydrodynamics.

% \citet{rettig2012} used a measured filament lengthscale and considered
% loss-limited and magnetic damping models separately for several historical
% SNRs.  For Tycho: they report $B = 310 \muG$ with $\Ecut = 24
% \unit{TeV}$; since $\Ecut$ is set by balancing advective/diffusive
% lengthscales, this is equivalent to our reporting of a diffusion coefficient
% (?).  With magnetic damping, they report $B = 150 \muG$ and a larger
% $\Ecut = 34 \unit{TeV}$.  Taken from \citet{rettig2012}: \citet{acciari2011}
% give $80$, $230 \muG$ for leptonic/hadronic models for gamma-ray
% emission in Tycho.

% eta2
% NOTE keep notes on eta2 < 1 updated (region 5 should be better, after I
% remove the bad FWHM)
The best-fit values of $\eta_2$, for both loss-limited and damped fits,
are very poorly constrained.  In the loss-limited fits, we observe no clear
trend of $\eta_2$ with $\mu$ (Table~\ref{tab:fits-loss}) excepting a general
preference for smaller $\eta_2$ at $\mu > 1$.  Loss-limited fits are generally
consistent with the Bohm limit on diffusion, favoring $\eta_2 \geq 1$; only
Regions 5 and 18 have best-fit $\eta_2$ values strictly below $\eta_2 = 1$
within error.  Damped fits permit smaller values of $\eta_2$ in contrast to
loss-limited models, perhaps as stronger damping and smaller $\eta_2$ have
cancelling effects on rim widths (recalling that in the damped model, smaller
$\eta_2$ gives larger rim widths).

% A quick discussion of mu
Our fits at multiple values of $\mu$ are not especially insightful.
The $\chi^2$ values for individual fits are variable and quite large, so we are
unable to favor or disfavor fits at different $\mu$ values.  But, neglecting
caution, we observed that values of $\mu \geq 1$ are qualitatively favored by
$\chi^2$ in almost all regions.  This trend may be an partially an artifact of
the correlation between $B_0$ and $\eta_2$ as they are not entirely independent
parameters, but they are less correlated for larger $\mu$; it may be said that
fits at smaller $\mu$ may have fewer (non-integer) degrees of freedom,
partially offsetting our observations.  \citetalias{ressler2014} also found
better model fits for $\mu \geq 1$ in the remnant of SN 1006 and noted that
\citet{reynolds2004} favored $\mu \geq 1$ on grounds of better modeled remnant
morphology, but it remains uncertain whether $\mu \geq 1$ diffusion is truly
preferred.

We do not explore $\mu$ dependence extensively in the damped model, but we find
that $\mu > 1$ permits smaller $B_0$ values and \emph{larger} $\eta_2$ values,
in keeping with our previous discussion of the effects of $B_0$ and $\eta_2$ on
rim width in both loss-limited and damped models.

\subsubsection{Tycho distance estimates}
% TODO need to update for damping results!!!!!

In our modeling we adopted a distance $d$ to Tycho of $3 \unit{kpc}$, but
various estimates for Tycho's distance range between $2.3$--$4 \unit{kpc}$
\citep{hayato2010}.  A larger remnant distance would increase both physical
filament widths and shock velocity estimates from proper motion (as used in our
models).

% lad/ldiff scalings give eta/B0 scalings
In the loss-limited model, we find roughly that $\eta_2 \propto d^2$ while
$B_0$ is approximately constant.  The advective lengthscale,
equation~\eqref{eq:lad}, may be rearranged to obtain:
\begin{align}
    B_0 =\; &(317 \muG) \left(\frac{v_d}{10^8 \unit{cm/s}}\right)^{2/3}
                \nonumber \\
            &\times \left(\frac{l_{\mt{ad}}}{0.01 \unit{kpc}}\right)^{-2/3}
                \left(\frac{h\nu}{1 \unit{keV}}\right)^{-1/3}
\end{align}
% TODO if we keep this equation, double check the 317 \muG carefully...
or, more simply,
\begin{equation}
    B_0 \propto \left(v_d\right)^{2/3}
                \left(l_{\mt{ad}}\right)^{-2/3} \nu^{-1/3} .
\end{equation}
Both $l_{\mt{ad}}$ and $v_d$ scale linearly with remnant distance $d$ and thus
their effects cancel in determining the magnetic field.  If diffusion is the
primary control on filament lengthscales, equation~\eqref{eq:ldiff} yields:
\begin{equation}
    \eta \propto \left(l_{\mt{diff}}\right)^2 B_0^{3} \nu^{-(\mu - 1)/2}
\end{equation}
Recall that $\eta \propto \eta_2$ since $E_h = E_2$ is constant for a given
$B_0$, from equation~\eqref{eq:diffcoeff}.  Similar results, primarily for
magnetic field $B_0$, were previously given by \citet{parizot2006}.

In practice, both advection and diffusion are important to our filament widths,
but both scalings $\eta_2 \propto d^2$ and $B_0$ constant occur in model fits
with varying distance.  When comparing full model fits with remnant distances
of $3 \unit{kpc}$ and $4 \unit{kpc}$, the deviation from the idealized scaling
is $\lesssim 1 \%$ for $B_0$ and $\abt 1$--$5\%$ for $\eta_2$.  As varying
remnant distance $d$ leaves width-energy scaling $\mE$ invariant, the relative
contributions of $l_{\mt{ad}}$ and $l_{\mt{diff}}$ should also be invariant.
Then both lengthscales should scale simultaneously with $d$, yielding the
observed behavior.
\note[Aaron]{need a logic check here}

% Damping and remnant distance
In the damped model, a larger distance $d$ will require larger damping lengths
from magnetic turbulence; our values of $a_b$ should be unchanged, however, as
$a_b$ here is given in units of shock radius $r_s$.
% TODO run more fits with varying remnant distance...
% After I've gotten new sets of numbers set up

% -------------------------------------------------------------------------
% Main null result -- cannot disfavor either damping or loss-limited models
% -------------------------------------------------------------------------
\subsection{Damping is poorly-constrained by X-ray width-energy dependence}

\begin{table*}[ht]
    \scriptsize
    \centering
    \caption{$m_\mt{E}$ from best model fits, loss-limited and damped
        \label{tab:fits-mE}}
    \input{suppl-fits-mE.tex}
    \tablecomments{Best fits correspond to those given in
        Table~\ref{tab:fits-all}}
\end{table*}

Width-energy dependence is not as sensitive a discriminant on damping versus
loss-limited models as may be intuitively expected.  Only in cases of very
sharp energy drop-off (data and best loss-limited fit $\mE \sim -0.5$) is
damping significantly disfavored as a control on rim widths (Region 18).  To
illustrate this behavior in our fits more clearly, Table~\ref{tab:fits-mE}
lists $\mE$ values as computed from both best loss-limited and damping fits.
This condenses and quantifies the width-energy predictions shown in
Figure~\ref{fig:fits}.

Both loss-limited and damped fits do not perfectly capture sharp drop-offs
(especially between $0.7$--$1 \unit{keV}$ and $1$--$2 \unit{keV}$), and the
data are greatly scattered.  In most regions, best fit width-energy curves have
$\mE \sim -0.2$ (Table~\ref{tab:fits-mE}).

% Explain the spurious energy dependence thing
When width-energy dependence is small, both loss-limited and damped rims give
best fits with very similar shapes -- strong fields at the shock in both models
give very sharp drop-offs in rim intensity.  When energy dependence is larger,
damping profiles can only yield energy dependent rims if magnetic fields are
fairly small (say, $\lesssim 50 \muG$) and rim half-maximums are achieved some
distance away from the damping lengthscale.  Away from the shock, magnetic
damping has already weakened and electron losses are controlled by synchrotron
losses which act differentially on the range of electron energies we consider.
% TODO horrific wording.  "act differentially?"

% TODO somehow highlight this, or think of a way to differentiate the regions
% Arbitrary yes, but I think it highlights the idea...
Each of these cases is illustrated in Figure~\ref{fig:rims}; in Region 1 the
damping profiles are quite distinct from the loss-limited profiles, whereas in
Region 16 they appear quite similar.  Regions 1, 2, 4, 5, 12, and 18 are best
fit in a damping model with fields $B_0 < 30 \muG$; the remaining regions have
$B_0 > 80 \muG$.  These regions are the same ones characterized by larger best
fit $\mE$ values in Table~\ref{tab:fits-mE}, with $\mE \lesssim -0.3$
generally, especially at lower energies.  The remaining regions are
characterized by $\mE \sim -0.2$, or $\mE$ decreasing (stronger width-energy
dependence) at \emph{higher} energies rather than lower energies.  The effect
of stronger energy dependence is conceptually expected in the damping model
when the advection lengthscale falls beneath the damping lengthscale at high
energies (Section~\ref{sec:energydep}).

% ------------------------------------------------------------------
% Next steps, observations of profiles (this might fit better above)
% ------------------------------------------------------------------
\subsection{Tactics for constraining diffusion}

\subsubsection{Fits to whole damping profiles}

The distinct profile shapes, of both thin rims and plateauing emission just
downstream of the thin rims, may be more sensitive discriminants of
loss-limited and damping models.  In particular, our best fits to data with
moderate $\mE$ generally require small magnetic field $B_0$ and comparatively
strong emission downstream of the thin rim, just below the rim half-maximum.
Although the FWHMs accorded with those of the loss-limited, the profiles were
quite different (Figure~\ref{fig:rims}).

Moreover, we selected only regions on Tycho where a FWHM could be resolved.
A subset of parameters (small damping lengths, small magnetic fields, and weak
or even sub-Bohm diffusion) can produce weaker thin rims that do not decrease
downstream to a half-maximum intensity; several nonthermal filaments around
Tycho exhibit similar behavior of a weak rim at the shock, followed by strong
downstream emission.  If profiles can be measured as far back as possible,
while avoiding thermal emission, we may be able to place stronger constraints
on allowed values of $B_0$, $\eta_2$, and $a_b$.

\subsubsection{Radio rim measurements}

Radio rims may provide an additional constraint on magnetic field structure.
The presence of radio rims may reflect 1. age-limited electron propagation (set
by the timescale needed to accelerate electrons and remnant expansion), 2.
possible reverse shock accelerated electrons, or 3. magnetic damping.  In our
steady-state transport model (equation~\eqref{eq:model}), the loss-limited
model \emph{cannot} produce radio rims.  Thin rims of radio emission in Tycho,
if not due to instabilities or other projected structure at the shock, can only
be attributed to magnetic damping.  But, the bulk of radio emission in Tycho
($r \lesssim 0.9 r_s$) may not be safely modeled as time-independent.

Synchrotron radio emission is emitted by \~ GeV electrons and hence diffusion
is essentially negligible; observed rims (or lack thereof) are primarily
controlled by $B_0$ and $a_b$.  It further follows that we needn't worry about
the precise electron cut-off energy, and we may be able to solve the transport
equation analytically (or otherwise circumvent the questionable assumption that
$D(x) B^2(x)$ is constant).  However, we note that our diffusion coefficients
have assumed that the mean free path is set by gyroradius (linearly so if
$\mu = 1$), and the Bohm limit $\eta = 1$ of our diffusion coefficients is
merely an expected lower limit on diffusion.  We may need a more accurate model
for particle diffusion at radio wavelengths.

The radio emission profiles from Tycho has been explored by
\citet{cassam-chenai2007} and \citet{morlino2012}; both studies assumed
moderate magnetic damping downstream of the shock.  Their discussions of radio
profiles were undertaken in context of a broader, more concerted hydrodynamic
modeling effort (which would be necessary to consider the internal radio
emission, rather than just the shock front emission).  In future work we will
consider a similar analysis as carried out in this paper, using radio rims to
help constrain X-ray rim parameters.

\subsubsection{Diffusion from synchrotron roll-off}

A major independent observable is the synchrotron roll-off frequency.  If
Bohm-like ($\mu = 1$) diffusion holds, the synchrotron roll-off is independent
of magnetic field strength and thus grants another constraint on diffusion
coefficient, within the assumption of DSA driven electron losses.  In this
manner, \citet{parizot2006} estimated diffusion coefficients $\eta = 10, 5$ for
compression ratios $4, 10$ in Tycho.

We can measure the roll-off frequency in all regions, fix the diffusion
coefficient at a given value of $\mu$, and obtain single parameter fits for
$B_0$ (or two parameter fits for $B_0$ and $a_b$) at each value of $\mu$.
Conceptually, this is comparable to our fits with $\eta_2 = 1$ fixed
(Table~\ref{tab:fits-all-eta2one}) and should not greatly affect our
conclusions.  However, if the diffusion coefficient from the roll-off frequency
is indeed well-constrained, then parameter estimates based on deduced diffusion
coefficients may be more reliable.

% -------------------------------------------
% Model correctness - are our results valid?!
% -------------------------------------------
\subsection{Model correctness}

% Data quality
We emphasize that our measurements are not data limited, as additional counts
from averaging measurements or selecting larger regions will not improve our
ability to constrain $B_0$ and $\eta_2$.  This is easily seen from
Figure~\ref{fig:fits} as well as large $\chi^2$ values in
Tables~\ref{tab:fits-loss} that reflect relatively tight errors in our FWHM
profile measurements (Table~\ref{tab:fwhms}).

The exponential variation in $D(x)$ (equation~\eqref{eq:ddamp}) may not be
physically reasonable; as we have noted above, the assumption $D(x)
\propto 1 / B^2(x)$ gives rise to sharp gradients in diffusion coefficient and
even inverts the effect of $\eta_2$ on our model profiles (where larger
$\eta_2$ can cause rim widths to narrow).

% TODO question: we have been ignoring intensity.  What if, e.g.,
% with strong diffusion particles get out of the damping zone faster and don't
% even have a chance to radiate, or something like that?
% I have no clue... check this, double check Sean's normalizations to be safe

%Expressions for $D$ in terms of unit power per log bandwidth, which is energy
%dependent as particles scatter preferentially on certain waves or whatever...
%could $D$ just be high at the shock (and that would also suggest taking $D
%\propto 1 / B_0$ rather than $D \propto 1 / \Bmin$, for our expression
%for cut-off energy $\Ecut$, could be a problem).

%Quickly remark on effect of changing compression ratio (effectively dropping
%$v_d$ downstream).  In both loss-limited and damped models, a higher
%compression ratio at the shock gives rise to a lower downstream velocity, which
%permits weaker magnetic fields and diffusion.

% ==========
% Conclusion
% ==========
\section{Conclusion}

We measured the widths of thin synchrotron filaments around Tycho's supernova
remnant and confirmed that our measured filaments are dominated by nonthermal
emission with clearly measurable full widths at half maximum in 4--5 energy
bands.  We find moderate narrowing of rim widths across the remnant,
corroborating rim narrowing observed by \citet{ressler2014} in the remnant of
SN 1006.  Our multiple measurements give loss-limited diffusion and magnetic fields
broadly consistent with prior estimates from rim widths
\citep[e.g.,][]{parizot2006, rettig2012} and radio and gamma ray measurements
\citep{acciari2011, morlino2012}.  We find that magnetic damping is equally
capable of describing our observed data.  At small energy dependence the two
models are indistinguishable and magnetic damping fits still favor moderately
amplified magnetic fields.  At moderate energy dependence ($\mE \sim -0.3$),
the damping model favors weaker magnetic fields, but modeled FWHMs may not
be truly representative of thin rim widths.

As our data are not count limited, we suggest that additional proxies for
diffusion and magnetic field are needed to establish or rule out the presence
of damping at the shocks of young supernova remnants.  More sophisticated
modeling of shock physics \citep[e.g.,][]{zirakashvili2014} and shock/filament
structure \citep[cf.][]{caprioli2013} may be necessary to improve modeling
results.  Complete rim profile fitting and subarcsecond radio profiles (as may
be enabled by EVLA and ALMA) may better discriminate between models.

% Further work -- radio
%As has been noted by many workers, observations of radio synchrotron rims would
%provide strong and immediate evidence for magnetic field damping (at the
%lengthscales of said radio rims).  Previous workers have certainly tried
%\citep{cassam-chenai2007, morlino2012} but it seems not easy.
%A new epoch of VLA and \Chandra observations in this year (PIs: J. Hewitt, B.
%Williams) will offer a rich dataset of sub-arcsecond observations that may
%reveal more about rim morphology and perhaps even temporal variation over a 10+
%year baseline (compare to \citet{katsuda2010-sn1006}).

% ================
% Acknowledgements
% ================
\acknowledgments

A.T. thanks CRESST for sponsorship/support at GSFC (any relevant grants?)
B.J.W. NPP?
The scientific results reported in this article are based on data obtained from
the \Chandra Data Archive.
This research has made extensive use of NASA's Astrophysics Data System.

{\it Facilities:} \facility{CXO (ACIS-I)}

% ========
% Appendix
% ========
\clearpage
\appendix

\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

% ==============================
% \mE values for best-fit curves
% ==============================
%\section{Best fit width-energy curve $m_\mt{E}$ values}
%
%\begin{table*}[ht]
%    \footnotesize
%    \centering
%    \caption{$m_\mt{E}$ computed from best loss-limited fit FWHMs}
%    \input{suppl-fits-mE.tex}
%\end{table*}

\section{Best fit stuff with $\Bmin=2\muG$}

\begin{table}[ht]
    \scriptsize
    \centering
    \caption{Best model fits for all regions, $\mu = 1$, $\Bmin=2\muG$}
    \input{suppl-fits-all-bmin2.tex}
\end{table}

% As above but with eta2=1 fixed!
\begin{table}[ht]
    \scriptsize
    \centering
    \caption{Best model fits for all regions, $\mu = \eta_2 = 1$,
        $\Bmin=2\muG$}
    \input{suppl-fits-all-bmin2-eta2one.tex}
\end{table}

% Only show these giant tables in 2col... no need to print these
\iftoggle{manuscript}{
}{
    % ===============================
    % Damping best fit sea of numbers
    % ===============================
    \clearpage
    \section{Magnetic damping best fit tables}

    \begin{table*}[ht]
        \scriptsize
        \centering
        \caption{Damped fits to individual regions, Filament 1, $\mu = 1$}
        \input{suppl-fits-damp-f1.tex}
    \end{table*}

    \begin{table*}[ht]
        \scriptsize
        \centering
        \caption{Damped fits to individual regions, Filament 2, $\mu = 1$}
        \input{suppl-fits-damp-f2.tex}
    \end{table*}

    \begin{table*}[ht]
        \scriptsize
        \centering
        \caption{Damped fits to individual regions, Filament 2 (cont.), $\mu = 1$}
        \input{suppl-fits-damp-f2-pt2.tex}
    \end{table*}

    \begin{table*}[ht]
        \scriptsize
        \centering
        \caption{Damped fits to individual regions, Filament 3, $\mu = 1$}
        \input{suppl-fits-damp-f3.tex}
    \end{table*}

    \begin{table*}[ht]
        \scriptsize
        \centering
        \caption{Damped fits to individual regions, Filament 4, $\mu = 1$}
        \input{suppl-fits-damp-f4.tex}
    \end{table*}

    \begin{table*}[ht]
        \scriptsize
        \centering
        \caption{Damped fits to individual regions, Filament 5, $\mu = 1$}
        \input{suppl-fits-damp-f5.tex}
    \end{table*}

    % ================================
    % Damping best fit tables, SN 1006
    % ================================
    \clearpage
    \section{Magnetic damping fit tables, SN 1006}

    \begin{table*}[ht]
        \scriptsize
        \centering
        \caption{Damped fits to SN1006 filaments, $\mu = 1$}
        \input{suppl-fits-damp-sn1006.tex}
    \end{table*}

    \clearpage
    \section{Magnetic damping fit tables, SN 1006, no e- cutoff}

    \begin{table*}[ht]
        \scriptsize
        \centering
        \caption{Damped fits to SN1006 filaments, $\mu = 1$, no e- cutoff}
        \input{suppl-fits-damp-nocut-sn1006.tex}
    \end{table*}
}

% =====================================================
% Synchrotron roll-off energy and diffusion coefficient
% =====================================================
\clearpage
\section{Derivation for synchrotron roll-off and diffusion}

% This is NOT optimized for 2-column layout

Assume that a DSA model explains the observed e-/synchrotron spectra
cut-off/roll-off.  We estimate e- cut-off energy by equating
$\tau_{\mt{accel}} = \tau_{\mt{synch}}$.  This gives (Parizot et al., 2006):
\begin{equation}
    \frac{3r}{r-1} \frac{r D_d + D_u}{v_s^2} = \frac{1}{b B^2 \Ecut}
\end{equation}
(\emph{I don't know if this equality was designed to exactly obtain the electron
cutoff energy, or is just a rough/scaling relation -- pull up Drury (1983) to
check this})
Following through the arguments/assumptions in Parizot et al. (2006) yields
(using $13.3 \unit{erg} = 8.3 \unit{TeV}$):
\begin{equation}
    \Ecut =
        \left( 13.3 \unit{erg} \right)^{\frac{2}{1+\mu}}
        \left( \frac{B_0}{100 \muG} \right)^{-\frac{1}{1+\mu}}
        \left( \frac{v_s}{10^8 \unit{cm\;s^{-1}}} \right)^{\frac{2}{1+\mu}}
        \eta^{-\frac{1}{1+\mu}}
\end{equation}
Now, we can convert this cut-off e- energy to a roll-off synchrotron energy by
invoking the delta function assumption $\nu_{\mt{cut}} = c_m \Ecut^2 B$ with
$c_m = 1.822 \times 10^{18}$.
(\emph{Again, I don't know if this is an equality or just a scaling.  We'd need
to relate the electron spectrum cut-off to the derived synchrotron spectrum
cut-off, as is done (I think) by Zirakashvili and Aharonian (2007).})
\begin{equation}
    \nu_{\mt{cut}} = c_m
        \left( 13.3 \unit{erg} \right)^{\frac{4}{1+\mu}}
        \left( \frac{B}{100\muG} \right)^{-\frac{1-\mu}{1+\mu}}
        \left( \frac{v_s}{10^8 \unit{cm/s}} \right)^{\frac{4}{1+\mu}}
        \eta^{-\frac{2}{1+\mu}}
\end{equation}
In the case of $\mu = 1$ this yields:
\begin{equation}
    \nu_{\mt{cut}} = (0.133 \unit{keV} / h)
        \left( \frac{v_s}{10^8 \unit{cm/s}} \right)^{2}
        \eta^{-1}
\end{equation}
For $\mu \neq 1$ this is no longer independent of magnetic field.  But,
there is hope.  Rewrite this in terms of $\eta_2$ using
$\eta = \eta_2 E_2^{1-\mu}$ and write:
\[
    \nu_{\mt{cut}} = c_m
        \left( 13.3 \unit{erg} \right)^{\frac{4}{1+\mu}}
        \left( E_2 \right)^{-\frac{2(1-\mu)}{1+\mu}}
        \left( \frac{B}{100\muG} \right)^{-\frac{1-\mu}{1+\mu}}
        \left( \frac{v_s}{10^8 \unit{cm/s}} \right)^{\frac{4}{1+\mu}}
        \left( \eta_2 \right)^{-\frac{2}{1+\mu}}
\]
With $E_2 = \left( (2 \unit{keV}/h) / (c_m B) \right)^{1/2}$, or
$E_2 = \left( 0.2657 \unit{erg^2\;G} / B \right)^{1/2}$, we obtain:
\[
    \nu_{\mt{cut}} = c_m
        \left( 13.3 \unit{erg} \right)^{\frac{4}{1+\mu}}
        \left( \frac{0.2657 \unit{erg^2\;G}}{B} \right)^{-\frac{1-\mu}{1+\mu}}
        \left( \frac{B}{100\muG} \right)^{-\frac{1-\mu}{1+\mu}}
        \left( \frac{v_s}{10^8 \unit{cm/s}} \right)^{\frac{4}{1+\mu}}
        \left( \eta_2 \right)^{-\frac{2}{1+\mu}}
\]
And, the magnetic field terms cancel!  Hey presto.
Our final usable result is:
\begin{equation}
    \nu_{\mt{cut}} = c_m
        \left( 13.3 \unit{erg} \right)^{\frac{4}{1+\mu}}
        \left( 0.2657 \unit{erg^2\;G} \right)^{-\frac{1-\mu}{1+\mu}}
        \left( 100 \muG \right)^{\frac{1-\mu}{1+\mu}}
        \left( \frac{v_s}{10^8 \unit{cm/s}} \right)^{\frac{4}{1+\mu}}
        \left( \eta_2 \right)^{-\frac{2}{1+\mu}}
\end{equation}

% --------------
% SN 1006 tables
% --------------
%\clearpage
%\section{Full model validation, SN 1006 (DRAFT ONLY)}
%
%For comparison to \citetalias{ressler2014}, we reproduce Sean's full model fit
%table (Table~\ref{tab:sean}) and present full model fits with 2 and 3 energy
%bands (Tables~\ref{tab:sn1006-2band} and \ref{tab:sn1006-3band}, respectively).
%We don't compare simple model results, as they are identical (only the error
%calculations differ).
%The procedures of Tables~\ref{tab:sean},~\ref{tab:sn1006-2band} are not quite
%the same.  Sean used $\mE$ at 2 keV instead of the $1$--$2$ keV width for
%fitting, whereas I used both widths.  But, it should not matter much as
%the fit often has $\chi^2 \sim 0$, either way (depending on the exact
%measurements and values of $\mu$).
%
%Please take the errors with a grain of salt.  They are automatically
%generated and \textbf{have not been manually validated}.  Some values may be
%invalid, where my error-finding code failed and gave a best, conservative guess
%of the error.
%
%\begin{table}[ht]
%    \tiny
%    \centering
%    \caption{Sean's SN 1006 best fit parameters \citepalias[Table 8]{ressler2014}.
%    \label{tab:sean}}
%    \input{tab-sean.tex}
%\end{table}
%
%\begin{table*}[ht]
%    \scriptsize
%    \centering
%    \caption{SN 1006 best fit parameters, 2 highest energy bands (full model).
%    \label{tab:sn1006-2band}}
%    \input{tab-sn1006-2band.tex}
%\end{table*}
%
%\begin{table*}[ht]
%    \scriptsize
%    \centering
%    \caption{SN 1006 best fit parameters, 3 energy bands (full model).
%    \label{tab:sn1006-3band}}
%    \input{tab-sn1006-3band.tex}
%\end{table*}

% ==========
% References
% ==========
\bibliographystyle{apj}  % AASTeX journal macros are supplied in ADS entries
\bibliography{refs-snr}

%\end{CJK}
\end{document}